---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sfdSAR

The goal of sfdSAR is to make it easy to follow the procedure of calculating swept
area ratio of an area of seabed by a fishing gear.

## Installation

You can install the sfdSAR from GitHub using

``` r
devtools::install_github("ices-tools-dev/sfdSAR")
```

## Example

The functions in this package are intended for one purpose: to compute the swept 
area ratio (SAR) and the subsurface SAR of a fishing gear, which can then be 
summarised over years and gear groupings.  

The code below shows how the sfdSAR functions can be used to calculate swept 
area ratio (SAR)

```{r example}
library(dplyr)
library(sfdSAR)

## load sample vms data
data(test_vms)

# load lookup tables
data(gear_widths)
data(metier_lookup)

# join widths and lookup
aux_lookup <-
  gear_widths %>%
  right_join(metier_lookup, by = c("benthis_met" = "Benthis_metiers"))

# add aux data to vms
vms <-
  aux_lookup %>%
  right_join(test_vms, by = c("LE_MET_level6", "LE_MET_level6"))

# calculate the gear width model
vms$gearWidth_model <-
  predict_gear_width(vms$gear_model, vms$gear_coefficient, vms)

# do the fillin for gear width:
# select provided average gear width, then modelled gear with, then benthis
# average if no kw or aol supplied
vms$gearWidth_filled <-
  with(vms,
    ifelse(!is.na(avg_gearWidth), avg_gearWidth / 1000,
      ifelse(!is.na(gearWidth_model), gearWidth_model / 1000,
        gearWidth)
    ))

# calculate surface contact
vms$surface <-
  predict_surface_contact(vms$contact_model,
                          vms$fishing_hours,
                          vms$gearWidth_filled,
                          vms$ICES_avg_fishing_speed)

# compute summaries over groups
output <-
  vms %>%
    mutate(
      mw_fishinghours = kw_fishinghours / 1000
    ) %>%
    group_by(year, c_square, Fishing_category_FO) %>%
    summarise(
      mw_fishinghours = sum(mw_fishinghours, na.rm = TRUE),
      subsurface = sum(surface * subsurface_prop * .01, na.rm = TRUE),
      surface = sum(surface, na.rm = TRUE)
    ) %>%
  ungroup %>%
  mutate(
    lat = sfdSAR::csquare_lat(c_square),
    lon = sfdSAR::csquare_lon(c_square)
  )

output
```

In this made up example, we have calculated the total MW fishing hours 
`mw_fishinghours`, subsurface swept area ratio (subsurface SAR) `subsurface` and 
surface swept area ratio (surface SAR) `surface`, for each Fishing category in 
a single c_square.

The above code can be applied to a larger dataset to covering a range of year, 
fishing gears and c_squares.

