---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sfdSAR

The goal of sfdSAR is to make it easy to follow the procedure of calculating swept
area ratio of an area of seabed by a fishing gear.

## Installation

You can install the sfdSAR from GitHub using

``` r
devtools::install_github("ices-tools-dev/sfdSAR")
```

## Example

The functions in this package are intended for one purpose: to compute the swept 
area ratio (SAR) and the subsurface SAR of a fishing gear, which can then be 
summarised over years and gear groupings.  

The code below shows how the sfdSAR functions can be used to calculate swept 
area ratio (SAR)

In the following examples the `dplyr` package is used to simplify the data 
processing and a made up vms toy vms dataset (`test_vms`) will be used
```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(sfdSAR)
data(test_vms)
```
```
library(dplyr)
library(sfdSAR)
## load sample vms data
data(test_vms)
```

### Predicting gear widths

The calculation of gear with is done using the data in the `gear_widths` table:

```{r gear_info, results="asis"}
data(gear_widths)
kableExtra::kable(gear_widths)
```

This table comes from (citation required), subsequently modified by (citation 
required).  And contains, for each benthis gear group, the proportion of the gear
contact that also affects the subsurface, the estimated average gear width,
and the coeffients and covariates of the surface contact model which relates
the gear width, properties of the vessel (kw or overall length) to 
bottom contact.

In order to use this data a lookup table is required linking Metier level 6 
codes to the benthis gear groupings listed above.  The lookup table is given in
the `metier_lookup` dataset and contains other gear groupings used in ICES
outputs.

```{r metier_lookup, results='asis'}
data(metier_lookup)
kableExtra::kable(head(metier_lookup))
```


Linking the gearwidths and contact model information is done with the following
two lines


```{r join}
# join widths and lookup
aux_lookup <-
  gear_widths %>%
  right_join(metier_lookup, by = c("benthis_met" = "Benthis_metiers"))

# add aux data to vms
vms <-
  aux_lookup %>%
  right_join(test_vms, by = c("LE_MET_level6", "LE_MET_level6"))
```

and the gear width model is applied using the helper function `predict_gear_width`

```{r gearwidth}
# calculate the gear width model
vms$gearWidth_model <-
  predict_gear_width(vms$gear_model, vms$gear_coefficient, vms)
```

In general, if gearwdth is available, it is used.  If average overall vessel 
length (oal) or average vessel power (kW) is available then the gear width
model is used.  FInally if none of these are avaiable an average gear width is 
applied.  The following code implements this

```{r fillin}
# do the fillin for gear width:
# select provided average gear width, then modelled gear with, then benthis
# average if no kw or aol supplied
vms$gearWidth_filled <-
  with(vms,
    ifelse(!is.na(avg_gearWidth), avg_gearWidth / 1000,
      ifelse(!is.na(gearWidth_model), gearWidth_model / 1000,
        gearWidth)
    ))
```

finaly, surface contact is computed using the appropriate surface contact
model, given by the `contact_model` feild, defined as:

```{r contact_model}
sapply(unique(gear_widths$contact_model), function(x) body(get(x)))
```

The helper function `predict_surface_contact` computes the surface contact 
(usage shown below).  The feild `subsurface_prop` which has come from the
`gear_width` dataset can be used to compute subsurface contact from the surface
contact.

```{r surface}
# calculate surface contact
vms$surface <-
  predict_surface_contact(vms$contact_model,
                          vms$fishing_hours,
                          vms$gearWidth_filled,
                          vms$ICES_avg_fishing_speed)
# calculate subsurface contact
vms$subsurface <- vms$surface * vms$subsurface_prop * .01
```


Normally it is required to summarise the surface quantities, which can be done
like this

```{r summary}
# compute summaries over groups
output <-
  vms %>%
    mutate(
      mw_fishinghours = kw_fishinghours / 1000
    ) %>%
    group_by(year, c_square, Fishing_category_FO) %>%
    summarise(
      mw_fishinghours = sum(mw_fishinghours, na.rm = TRUE),
      subsurface = sum(subsurface, na.rm = TRUE),
      surface = sum(surface, na.rm = TRUE)
    ) %>%
  ungroup %>%
  mutate(
    lat = sfdSAR::csquare_lat(c_square),
    lon = sfdSAR::csquare_lon(c_square)
  )

output
```

In this made up example, we have calculated the total MW fishing hours 
`mw_fishinghours`, subsurface swept area ratio (subsurface SAR) `subsurface` and 
surface swept area ratio (surface SAR) `surface`, for each Fishing category in 
a single c_square.

The above code can be applied to a larger dataset to covering a range of year, 
fishing gears and c_squares.


```{r ref.label=c("join", "gearwidth", "fillin", "surface", "summary")}

```
